Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude

### 2016-03-21 RMS/submission.tex
partial editting

Time affects the system by advancing both the running task whose $ID$
is loaded at \verb|pc| and the interrupt source simultaneously.  While
time elapses, \verb|cnt| of the former increases and \verb|val| of the
latter decreases, respectively:
\begin{verbatim}
  ceq delta((L T STS HW ISRC), R)
      = ((deltaTask(ID, L, R) 
         T STS HW (deltaIS(ISRC, R)))
      if ID := (HW).getPc /\ ID :: MaybeNat .
\end{verbatim}
We omit details for the case where \verb|ID| is of sort \verb|Oid| due
to similarity. In that case, \verb|deltaTask| applies on \verb|STS|
instead of \verb|L|.

@@

First we need to introduce some more preliminaries
from~\cite{DBLP:journals/entcs/OlveczkyM07a}. A term $t$ is called
\emph{ground} if it contains no variables. For a set $P\subseteq AP$
of atomic propositions and ground terms $t,t'$, we write $t\simeq_P
t'$ (or $t\simeq t'$ for simplicity when $P$ is implicit) if $t$ and
$t'$ satisfy exactly the same set of propositions from $P$.

The following lemma is needed to prove the time-robustness of our
model:
\begin{lemma}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
\label{l:timerobustness}
Let $\cal R$ be an object-oriented specification with a standard tick
rule, let the infinity element \verb|INF| be the only element in the
time domain which is not a normal time value, and let the time domain
be \emph{linear}. Then $\cal R$ is time-robust if the following
conditions are satisfied for all appropriate ground terms $t$ and
$r,r'$:

\noindent
(i) \verb|mte(delta(|$t$\verb|,|$r$\verb|))| $=$
\verb|mte(|$t$\verb|) monus |$r$, for all $r\le$
\verb|mte(|$t$\verb|)|;

\noindent
(ii) \verb|delta(|$t$\verb|,|$0$\verb|)| $= t$;

\noindent
(iii) \verb|delta(delta(|$t$\verb|,|$r$\verb|),|$r'$\verb|)| $=$
\verb|delta(|$t$\verb|,|$r+r'$\verb|)|, for $r+r'\le$
\verb|mte(|$t$\verb|)|;

\noindent
(iv) \verb|mte(|$\sigma(l)$\verb|)|$= 0$ for each ground instance
$\sigma(l)$ of each left-hand side of an instantaneous rewrite rule.
\end{lemma}

@@

Now we can present the detailed proof of Theorem~\ref{t:main}:
\noindent
\begin{IEEEproof}[Proof of Theorem~\ref{t:main}]
first
\end{IEEEproof}

>>>>

Time affects the system by advancing both the running task whose $ID$
is loaded at \verb|pc| and the interrupt source simultaneously.  While
time elapses, \verb|cnt| of the former increases and \verb|val| of the
latter decreases, respectively:
\begin{verbatim}
  ceq delta((L T STS HW ISRC), R)
      = (deltaTask(ID, L, R) 
         T STS HW (deltaIS(ISRC, R)))
      if ID := (HW).getPc /\ ID :: MaybeNat .
\end{verbatim}
We omit details for the case where \verb|ID| is of sort \verb|Oid| due
to similarity. In that case, \verb|deltaTask| applies on \verb|STS|
instead of \verb|L|.

@@

Some more preliminaries from~\cite{DBLP:journals/entcs/OlveczkyM07a}
are needed. A term $t$ is called \emph{ground} if it contains no
variables. For a set $P\subseteq AP$ of atomic propositions and ground
terms $t,t'$, we write $t\simeq_P t'$ (or $t\simeq t'$ for simplicity
when $P$ is implicit) if $t$ and $t'$ satisfy exactly the same set of
propositions from $P$.

The following lemma is needed to prove the time-robustness of our
model:
\begin{lemma}[\cite{DBLP:journals/entcs/OlveczkyM07a}]
\label{l:timerobustness}
Let $\cal R$ be an object-oriented specification with a standard tick
rule, let the infinity element \verb|INF| be the only element in the
time domain which is not a normal time value, and let the time domain
be \emph{linear}. Then $\cal R$ is time-robust if the following
conditions are satisfied for all appropriate ground terms $t$ and
$r,r'$:
\begin{enumerate}
\item
[(i)] \verb|mte(delta(|$t$\verb|,|$r$\verb|))| $=$
\verb|mte(|$t$\verb|) monus |$r$, for all $r\le$
\verb|mte(|$t$\verb|)|;

\item
[(ii)] \verb|delta(|$t$\verb|,|$0$\verb|)| $= t$;

\item
[(iii)] \verb|delta(delta(|$t$\verb|,|$r$\verb|),|$r'$\verb|)| $=$
\verb|delta(|$t$\verb|,|$r+r'$\verb|)|, for $r+r'\le$
\verb|mte(|$t$\verb|)|;

\item
[(iv)] \verb|mte(|$\sigma(l)$\verb|)|$= 0$ for each ground instance
$\sigma(l)$ of each left-hand side of an instantaneous rewrite rule.
\end{enumerate}
\end{lemma}

@@

Now we can present the detailed proof of Theorem~\ref{t:main}:
\noindent
\begin{IEEEproof}[Proof of Theorem~\ref{t:main}]
We first prove the time-robustness of our model by
Lemma~\ref{l:timerobustness}.  Instantiated on the built-in dense time
domain \verb|POSRAT-TIME-DOMAIN-WITH-INF|, our model has \verb|INF| as
the only element which is not a normal time value. As presented in
Section~\ref{ss:timedbehavior}, only a single standard tick rule is
defined. Hence our model is time-robust by
Lemma~\ref{l:timerobustness} provided conditions (i-iv) hold.
\begin{enumerate}
\item[(i)] We must prove \verb|mte(delta(|$s$\verb|,|$r$\verb|))| $=$
  \verb|mte(|$s$\verb|)| \verb|monus| $r$ for all $r\le$
  \verb|mte(|$s$\verb|)|, with $s$ being a system state of sort
  \verb|System|.
\end{enumerate}

\end{IEEEproof}
